---
title: 'Extra Tutorial: R Markdown'
author: "Semuhi Sinanoglu"
date: "September 2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
header-includes:
- \usepackage{placeins}
- \usepackage{ragged2e}
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)
library(tinytex) #use this in the setup chunk to knit to pdf if you do not have a tex installation. See above note.
#Put the packages that you'll use here. 

minwage <- read.csv("minwage.csv") 
```

# Q1

## Answer to 1.1 (subheading)

```{r code1, eval=TRUE, results='hide', echo=TRUE, message=FALSE}
#eval: to run the code or not
#you can suppress the results of this chunk or show it as it is, or you write 'hide'
#you can hide the code in your final output through echo
#suppress messages automatically displayed when you run a code like acknowledgements

schools <- expand.grid(status=c("Public", "Private"), 
                       funding=seq(1500, 2000, by=100), 
                       teacher=c(seq(5,15,by=5),NA)) 

mean(schools$teacher, na.rm = TRUE)

```

# Q2

You can drop your comments here. For a new line:\

You can write mathematical equations like this: $20 \geq \alpha \geq 76$.\
You can make **things bold** or *italic*.

For hyperlinks: [R project](https://www.r-project.org/)

If you want to leave some larger gap:

$~$

You can itemize:

-   Like this...
-   And this...
-   and goes on...

Another kind of listing:

1. Coding is fun.
2. Yes, I mean it.
3. Hang in there!

# Q3

## How to create a table

There are different ways in which you can create tables. If you want to create a table of an existing dataset, use the kable function.

I will not show you all the details, but you can put multiple tables side by side. Literally change anything with the outlook of your table with kable and kableExtra packages.

```{r table, echo=FALSE}
knitr::kable(minwage[1:10,], 
             caption = "Minimum wage dataset",
             align="lccccccc", booktabs=TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position"), font_size = 10) #hold position to avoid floating
```

\newpage


 Name                 Description
 -------------------- ---------------------------------------------------------------------------------
 `election_id`        Unique identifier for each election
 `country`            
 `year`               
 `party_name`         Name of the political party translated to English
 `vote_share`         Votes for the party as a percentage of total cast
 `seats`              The number of seats the party won
 `seats_total`        Total seats obtained in an election (for all parties)
 `left_right`         Average left-right ideology from a variety of sources (higher = more right)
 `state_market`       Average economic left-right score from two sources (higher = more right)
 `liberty_authority`  Average social left-right score from two sources (higher = more right)
 `party_family`        Party family 
 `halfdecade`         Five-year intervals to make aggregation easier


| **Date**  | **Cases** |
|-----------|-----------|
| 30/8/2020 |   371     |
| 29/8/2020 |   350     |
| 28/8/2020 |   300     |
| 27/8/2020 |   267     |
| 26/8/2020 |   226     |
| 25/8/2020 |   192     |
| 24/8/2020 |   174     |
\centering
**Table: COVID cases in Toronto**
\justify

You can set figure width and height:
\FloatBarrier
```{r, fig.width = 4, fig.height= 3, fig.align = 'center'}
plot(density(rnorm(1000)), main = "1000 Draws from a Standard Normal")
```
