---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(here)
afghan <- read.csv("afghan.csv")
knitr::opts_chunk$set(echo = FALSE)
```

Summarize variables of interest
```{r}
summary(afghan$age)
summary(afghan$educ.years)
summary(afghan$employed)
afghan$income <- as.factor(afghan$income)
summary(afghan$income)

```

Analyze violence experienced by ISAF and Taliban
```{r}
prop.table(table(ISAF = afghan$violent.exp.ISAF, Taliban = afghan$violent.exp.taliban))
prop.table(table(ISAF = afghan$violent.exp.ISAF, Taliban = afghan$violent.exp.taliban, exclude = NULL))
```


```{r}
ISAF.tb <- tibble(table(ISAF = afghan$violent.exp.ISAF))
ISAF.tb
```

Make a bar plot for ISAF
```{r}
ISAF.pt <- prop.table(table(ISAF = afghan$violent.exp.ISAF, exclude = NULL))
barplot( ISAF.pt, 
         names.arg = c("no harm", "harm", "nonresponse"), 
         main = "Civilian victimization by ISAF",
         xlab = "Response Category",
         ylab = "proportion of respondents", ylim = c(0, 0.7))


```

Make a bar plot for Taliban
```{r}
Taliban.pt <- prop.table(table(Taliban = afghan$violent.exp.taliban, exclude = NULL))
barplot( Taliban.pt, 
         names.arg = c("no harm", "harm", "nonresponse"), 
         main = "Civilian victimization by Taliban",
         xlab = "Response Category",
         ylab = "proportion of respondents", ylim = c(0, 0.7))
```

Create a histogram to measure the density distribution of the age category

```{r}
hist(afghan$age, freq = FALSE, ylim = c(0, 0.04), xlab = "age", main = "Distribution of respondent's age")
```

make a histogram for aftghan education years

```{r}
hist(afghan$educ.years, freq = FALSE, 
     breaks = seq(from = -0.5, to = 18.5, by = 1),
     xlab = "Years of education", 
     main = "Distribution of respondent's education")
##add a text label at (x, y) = (3, 0.5)
text(x = 3, y = 0.5, "median")
##add a vertical line representing median
abline(v = median(afghan$educ.years))

```

making a boxplot

```{r}
boxplot(afghan$age, main = "Distribution of age", ylab = "age", ylim = c(10, 80))
```

box plot with multiple variables

```{r}
boxplot(educ.years ~ province, data = afghan,
        main = "education by province", ylab = "years of education")
```

how does this compare to violence experienced by province?

```{r}
tapply(afghan$violent.exp.taliban, afghan$province, mean, na.rm = TRUE)
tapply(afghan$violent.exp.ISAF, afghan$province, mean, na.rm = TRUE)
```

sample validation 
```{r}
##load village data
afghan.village <- read_csv("afghan-village.csv")
##boxplots for altitude
boxplot(altitude ~ village.surveyed, data = afghan.village, 
        ylab = "Altitude (m)", names = c("nonsampled", "sampled"))
boxplot(log(population) ~ village.surveyed, data = afghan.village, 
        ylab = "log.population", names = c("nonsampled", "sampled"))
```

calculating non-response rates

```{r}
tapply(is.na(afghan$violent.exp.taliban), afghan$province, mean)
tapply(is.na(afghan$violent.exp.ISAF), afghan$province, mean)
```

list response for sensitive items

```{r}
mean(afghan$list.response[afghan$list.group == "ISAF"]) - mean(afghan$list.response[afghan$list.group == "control"])
```

floor and ceiling effects
```{r}
table(response = afghan$list.response, group = afghan$list.group)
```






