---
title: "Problem Set 2 (Sept 29)"
author: "Jamie Duncan"
date: "23/09/2020"
header-includes:
  - \usepackage{placeins}
  - \usepackage{ragged2e}
urlcolor: blue
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(knitr)
library(kableExtra)
library(tinytex)

boston <- read_csv("boston.csv")
prop.AJPS <- read_csv("proposalAJPS.csv")
```
# Section 1
## Question 1
### The benefit of randomly assigning individuals to the treatment or control groups is that the two groups should be similar, on average, in terms of their covariates. This is referred to as covariate balance. 

### 1.1.1: Show that the treatment and control groups are balanced with respect to the income variable (income) by comparing its distribution between those in the treatment group and those in the control group.

```{r echo=TRUE}
summary(as.factor(boston$income)) # looking at the data
summary(boston$income)
mean(boston$male)
```

```{r echo=TRUE}
## subset into treated and not treated
treated <- subset(boston, subset = (treatment == 1))
not_treated <- subset(boston, subset = (treatment == 0))

```

```{r echo=TRUE}
summary(treated$income) #summary of both treated then not treated
summary(not_treated$income)
```

```{r echo=TRUE}
median(treated$income) # median income for treated pop
median(not_treated$income)  #median income for not treated

median(treated$income) - median(not_treated$income) #subtract one from the other

```

In this case, median is the best single number to compare income because it is less sensitive to outliers (though looking at all of the numbers is important to see how direction outliers skew the data). It is perfectly even between te treated and non treated groups.

### 1.1.2: Also, compare the proportion of males (male) in the treatment and control groups. 

```{r echo=TRUE, warning=FALSE}
treated.men <- data.frame(prop.table(table(male = boston$male, treatment = boston$treatment))) # creating a proportion table to see the distribution of genders across treated and non-treated samples.
treated.men
```

```{r echo=TRUE, warning=FALSE}
treated.men[4, ] - treated.men[2, ] #calculating the difference between treated males and non-treated males
```
A proportion table helps us compare the proportion of treated versus non-treated men relative to the whole data set. We see that treated men account for roughly 9% less of the population included in the data than non-treated men.

### 1.1.3: Interpret these two numbers.
- The distribution is accepable.
- The median income of the treated and non treated participants is perfectly even. On average, it appears higher earning people were less likely to be included.
- Men were more likely not to be treated than treated. While imbalanced in that sense, by treating fewer men, there is better gender balance in the treated sample between men and non-male identified participants.

## Question 2
### Individuals in the experiment were asked a series of questions both at the beginning and the end of the experiment. One such question was “Do you think the number of immigrants from Mexico who are permitted to come to the United States to live should be increased, left the same, or decreased?” The response to this question prior to the experiment is in the variable numberim.pre. The response to this question after the experiment is in the variable numberim.post. In both cases the variable is coded on a 1 – 5 scale. Responses with values of 1 are inclusionary (pro-immigration) and responses with values of 5 are exclusionary (anti-immigration). 

### 1.2.1: Compute the average treatment effect on the change in attitudes about immigration. That is, how does the mean change in attitudes about immigration policy for those in the control group compare to those in the treatment group.

```{r echo=TRUE, warning=FALSE}
## writing a function for SATE
sateA <- function(df){ # it takes a data frame as its argument
treated <- subset(df, subset = (treatment ==1 ))# splits into treated/ non-treated subsets
not_treated <- subset(df, subset = (treatment == 0))
treated.shift <- (mean(treated$numberim.post - treated$numberim.pre, na.rm = TRUE)) #cacculates the average shift in attitudes towards immigration for both
nottreated.shift <- (mean(not_treated$numberim.post - not_treated$numberim.pre, na.rm = TRUE)) 
sate <- treated.shift - nottreated.shift #calculates the difference between the treated shift and the non treated shift
return(list(treated.shift, nottreated.shift, sate)) #returns three numbers: treated shift, non-treated shift, and the SATE
}
sateA(boston)


```

### 1.2.2: Interpret the result.
The participants treated with hispanic confederates on the train platform shifted their views on immigration policy slightly towards exclusion, whereas those not treated expressed more inclusionary tendencies after the 10 days. This contributed to a SATE of roughly 0.31, which means that the treated group became more xenophobic through exposure to hispanic confederates.

## Question 3
### Does having attended college influence the effect of being exposed to outsiders on exclusionary attitudes? Another way to ask the same question is this: is there evidence of a differential impact of treatment, conditional on attending college versus not attending college? 

### 1.3.1: Calculate the necessary quantities to answer this question and interpret the results. 

```{r echo=TRUE}

##subsetting data for college vs non-college
college <- subset(boston, boston$college == 1)
no.college <- subset(boston, boston$college == 0)

sateB <- function(df){ # this function is the same as sateA except it only returns the SATE figure
treated <- subset(df, subset = (treatment ==1 ))
not_treated <- subset(df, subset = (treatment == 0))
treated.shift <- (mean(treated$numberim.post - treated$numberim.pre, na.rm = TRUE))
nottreated.shift <- (mean(not_treated$numberim.post - not_treated$numberim.pre, na.rm = TRUE))
sate <- treated.shift - nottreated.shift
return(sate)
}

sate.coll <- sateB(college) #calculating SATE for both the collage and no college subsets
sate.nocoll <- sateB(no.college)

sate.coll - sate.nocoll #calculating the difference between them


```

There is a slight tendency towards exclusionary views of roughly 0.02 points among those participants that are college educated, which I interpret as a negligible difference.

## Question 4

### Repeat the same analysis as in the previous question but this time with respect to age and ideology.

### 1.4.1: For age, divide the data based on its quartile and compute the average treatment effect within each of the resulting four groups.

```{r echo=TRUE}
quantile(boston$age) #determining the boundaries of the age groups
```

```{r echo=TRUE}
boston$age_grp <- boston$age # creating a new column for age groups and inserting the age variable as a placeholder
boston$age_grp[boston$age_grp >0 & boston$age_grp <= 33] <- "20-33" # using logical statements to place the ages in the age group column into 4 categories
boston$age_grp[boston$age_grp >= 34 & boston$age_grp <= 43] <- "34-43"
boston$age_grp[boston$age_grp >= 44 & boston$age_grp <= 52] <- "44-52"
boston$age_grp[boston$age_grp >= 53] <- "53+"

glimpse(boston)#using glimpse to verify that the above worked as anticipated

```

```{r echo=TRUE}
##creating subsets for each of the newly formed age groups
ag1 <- subset(boston, boston$age_grp == "20-33")
ag2 <- subset(boston, boston$age_grp == "34-43")
ag3 <- subset(boston, boston$age_grp == "44-52")
ag4 <- subset(boston, boston$age_grp == "53+")

##running the same sate function as above on each age group ans presenting them in a date frame for ease of interpretation
data.frame("20 to 33" = sateB(ag1),"34 to 43" = sateB(ag2), "44 to 52" = sateB(ag3),"53 plus" = sateB(ag4))
```

### 1.4.2: For ideology, compute the average treatment effect within each value. 

```{r echo=TRUE}
##creating subsets for each ideology score
id1 <- subset(boston, boston$ideology == 1)
id2 <- subset(boston, boston$ideology == 2)
id3 <- subset(boston, boston$ideology == 3)
id4 <- subset(boston, boston$ideology == 4)
id5 <- subset(boston, boston$ideology == 5)

data.frame("1" = sateB(id1), "2" = sateB(id2), "3" = sateB(id3), "4" = sateB(id4), "5" = sateB(id5))

```

### 1.4.3: What patterns do you observe? Give a brief substantive interpretation of the results.
- Across the age groups, we can see that the SATE is positive for all but the 44-52 age group, indicating that each of the other age groups saw an exclusionary shift in their views on immigration. The shift is most pronounced in the 20-33 age group, which has a SATE of roughly 0.7 and the 53 plus group which has a SATE of 0.5.
- Among the ideology subsets, we see a curious trend whereby the biggest shifts towards exclusion (relative to the non treated group) are among those that would be described as "Liberal" (ideology = 2) with a SATE of 0.8 and those who are "Very Conservative" (ideology = 5) with a SATE of 0.75. Those who identify as "Moderates" (ideology = 3) were the only group that saw negative SATE meaning the treated subset shift tended toward inclusion more than the non-treated subset.

# Section 2
## Question 1
### 2.1.1: How many members of parliament are in the data?
```{r echo=TRUE, warning=FALSE}

##the data is done by MP so the length of any one full column will be equal to the number of MPs in the data
length(prop.AJPS$election)
```
### 2.1.2: Why is it not simply 338 x 2 = 676?

The paper sampled from *incumbent* MPs. Also the number of seats in the House of Commons is constantly shifting; there are 336 sitting MPs right now according to ourcommons.ca

### 2.1.3: What proportion of the observations in the data are members of the government party?

```{r echo=TRUE}
prop.table(table(prop.AJPS$gov))

mean(prop.AJPS$gov)
```

25% of the observations are from members of the governing party.

### 2.1.4: What proportion are female? 
```{r echo=TRUE}
prop.table(table(prop.AJPS$female))
```

About 20% of the observations are female MPs.

### 2.1.5: How much does a typical member spend on reelection?

```{r echo=TRUE}
mean(prop.AJPS$spending[prop.AJPS$spending >= 0]) #calculated mean without all the negative numbers that would erroneously skew the results

```

The average re-election spend is 63349.94.

### 2.1.6: Is there anything about the spending data that is odd?
```{r echo=TRUE}
summary(prop.AJPS$spending[prop.AJPS$spending < 0 ]) #seeing if there is a range of negative numbers or only one.
length(prop.AJPS$spending[prop.AJPS$spending < 0 ]) #seeing how many entries there are
```

There are 29 entries of -99999, which I omitted from calculations as if they were null entries.

## Question 2
### First, we will check to see if the randomization procedure appears to have been fair. Calculate the difference between the proportion of MPs who were able to propose legislation under this procedure according to whether they are female. That is, a government MP, and whether their vote share in the previous election was greater than or less than the median.

Semuhi's instructions to Hayley as posted in Slack: Calculate the difference between the proportion of MPs who were able to propose legislation under this procedure according to 1) whether they are female, 2) whether they are a government MP, and 3) whether their vote share in the previous election was greater than or less than the median.

This question was confusing but I calculated the difference between the treated group and the non-treated group (i.e. power to propose and no power to propose.) I know some others just dropped the untreated data set and calculated the difference between female/not female, gov/not gov for the treated sample.


```{r echo=TRUE}
## wrote a function to calculate the the difference between the proportion of MPs who were able to propose legislation
rand.prop <- function (df){ #rand.prop takes a data frame (df) as an argument
 p2p1 <- subset(df, p2p == 1)#it subsets that df into those with and without the power to propose
 p2p0 <- subset(df, p2p == 0)
 prop.fem <- mean(p2p1$female) - mean(p2p0$female) #it calculates the difference between the average proportion of females in each subset
 prop.gov <- mean(p2p1$gov) - mean(p2p0$gov) #it calculates the difference between the average proportion of government MPs in each subset
 median.prev <- median(df$previous_vote) #it calculates the median vote share for the previous vote
 prop.vs <- (median(p2p1$previous_vote) - median.prev) - (median(p2p0$previous_vote) -  median.prev) # it calculates the difference between the median vote share of each subset and the median overall and then the difference between the treated difference and the non treated difference. 
 out <- list(prop.fem, prop.gov, prop.vs) # the desired outputs are placed in a list
 names <- c("fem diff", "gov diff", "median diff")#the outputs are named
 names(out) <- names
 return(out) #they are then returned
}
rand.prop(prop.AJPS)

```
## Question 3

### 2.3.1: Calculate the mean vote share for MPs able to introduce legislation under this procedure and those unable to do so. 

```{r echo=TRUE}

##creating a new column for vote share that amounts to the average between the previous and current vote shares 
prop.AJPS$vote_share <- (prop.AJPS$previous_vote + prop.AJPS$current_vote) / 2

##using aggregate to calculate vote share indexed by power to propose.
vs.by.p2p <- aggregate(prop.AJPS$vote_share, list(p2p = prop.AJPS$p2p), mean)
vs.by.p2p


```

### 2.3.2: How big is the difference?

```{r echo=TRUE}

##subtracting the treated (p2p = TRUE) from the non-treated (p2p = FALSE)
vs.by.p2p[2, 2] - vs.by.p2p[1, 2]

```

The difference is about 1%

### 2.3.3: Create a table of the power to propose and whether the MP won reelection.

```{r echo=TRUE}
## Created a prop.table of with winners and p2p and converted it to a data frame to make it easier to read

data.frame(prop.table(table(winner = prop.AJPS$winner, p2p = prop.AJPS$p2p)))

```

### 2.3.4: Did the power to propose help MPs win reelection? 
It appears that MPs *without* the power to propose were more likely to win reelection by about 17 percent.

## Question 4

### 2.4.1: Create two new data frames, one with just MPs in the governing party and one with just members of the opposition. 

```{r echo=TRUE}
gov <- data.frame(subset(prop.AJPS, gov == 1))
opp <- data.frame(subset(prop.AJPS, gov == 0))

head(gov)
head(opp)
```
### 2.4.2: Now repeat what you did for Question 2 on each of these data frames. 

```{r echo=TRUE}
rand.prop(gov)

```

```{r echo=TRUE}
rand.prop(opp)

```

### 2.4.3: Is there a big difference in vote share or win share across the power to propose in either data frame? Do the differences look big?'

```{r echo=TRUE}
## writing a function so I don't have to do this twice
vote.win.share <- function (df){
 p2p1 <- subset(df, p2p == 1)#subset according to power to propose (yes/no)
 p2p0 <- subset(df, p2p == 0)
 vs1 <- mean(p2p1$vote_share)#calculating the mean vote share for each
 vs2 <- mean(p2p0$vote_share)
 ws1 <- mean(p2p1$winner)#calculating the mean win share for each
 ws2 <- mean(p2p0$winner)
 vsdiff <- vs1 - vs2
 wsdiff <- ws1 - ws2
 out <- c(vsdiff, wsdiff)
 names <- list("vote share", "win share")
 names(out) <- names
 return(out)
}

vote.win.share(gov)
vote.win.share(opp)

 

```

Most of the differences hover around zero but it appears that government MPs with the power to propose get, on average, 4.5 more of the vote share compared to those who don't.

## Question 5
### One possibility is that government MPs did better and opposition MPs did worse when they got proposal power because they draw lower quality opponents (as measured by previously holding electoral office). 

### 2.5.1: Did MPs with proposal power draw lower quality opponents in the next election? 

```{r echo=TRUE}
tapply(prop.AJPS$quality, prop.AJPS$p2p, mean) #comparing the average quality of opponent by power to propose

```

It doesn't appear so but I tried to look at this in another way:

```{r echo=TRUE}
prop.table(table(quality = prop.AJPS$quality, p2p = prop.AJPS$p2p))
```

Again, not seeing any huge indicators that power to propose is correlated with higher or lower quality opponents.

### 2.5.2: Did this vary by whether they were in the government party?
```{r echo=TRUE}
## performing same calculation as in 2.5.1 for the gov and opp subsets and then calculating the differences
tapply(gov$quality, gov$p2p, mean) - tapply(opp$quality, opp$p2p, mean)
```

It doesn't appear to vary by whether they are in the governing party either.

### Another possibility is that proposal power shaped the MPs ability to fundraise, by letting them appear to be bringing home the pork. 

### 2.5.3: Did proposal power lead MPs to spend more or less money? Rather than answering this with just a difference in means, create a boxplot for each of four groups based on the 2x2 of government vs. opposition and proposal power and no proposal power. 

```{r echo=TRUE}
## cleaning up the spending data to remove those pesky negative values.
spend.clean <- subset(prop.AJPS, spending >= 0) 
##exploring the data a bit before I jump in.
summary(spend.clean$spending)
summary(prop.AJPS$spending)
```


```{r echo=TRUE, warning=FALSE}
par(mfrow=c(1,2)) # I want to see the plots next to each other (1 row, 2 columns)
plot1 <- boxplot(log(spending) ~ p2p, spend.clean, #'logging' the spending axis and creating a boxplot comparing p2p in the newly cleaned data
 ylab = "Spending(log)", names = c("Prop", "No Prop"), #labelling elements of the plot
 ylim = c(10,11.5), na.rm = T) #setting the boundaries of the y-axis and removing null values
plot2 <- boxplot(log(spending) ~ gov, spend.clean, #same as above but for the gov variable
 ylab = "Spending (log)", names = c("Gov", "Non-gov"),
 ylim = c(10, 11.5), na.rm = T)

```

### 2.5.4: Provide a clear description as if it is in a paper (“This figure shows that. . . ”)

This figure shows that, despite some outliers, median spending for reelection across MPs in the government party and opposing parties with and without the power to propose is similarly distributed.